Argennon: A Scalable Smart Contract Platform

### 2022-04-15 tex/sections/avm.tex
Improve readability of "Resource Management" subsection
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

An Argennon Standard Representation (ASR) is a programming language for describing argennon applications, optimized
for the architecture and requirements of the Argennon platform.
Argennon supports two standard representations: one is a high level text based language which needs
costly compilation before being executed on a hardware machine. The other is a low level binary representation which
usually can be executed, with minimal pre-processing, by a JIT compiler or an emulator. The high level
language is intended for preserving the high level information of applications logic to facilitate
platform specific compiler optimization at host nodes. The low level language, is designed for
efficient direct execution of applications that are not frequently used.

The state of an Argennon application is stored in byte addressable finite arrays of memory called
\emph{heap chunks}. An application may have several heap chunks with different sizes, and can remove or
resize its heap chunks or allocate new chunks. Every chunk belongs to exactly one application and can only be modified
by its owner. In addition to heap chunks, every application also has an amount of non-persistent local memory for
storing temporary data.

@@

The Argennon Smart Contract Execution Environment can be seen as a machine for executing Argennon applications to
fulfill HTTP requests, produce their HTTP responses and update related heap chunks. The AscEE executes requests
sequentially\footnote{Actually requests are executed in parallel but by performing data dependency analysis the
result is guaranteed to be identical
with sequential execution of requests.} and each request has a separate \emph{execution session}. Execution sessions
are separate sessions of executing smart contract's code in order to fulfill \emph{external} HTTP requests. An
external request is a request that is not made by another Argennon application.

@@

During an execution session an application can make \emph{internal} HTTP requests to other applications. Those
requests will not start a new execution session and will be executed within the current session. In AscEE making an
HTTP request to an application is similar to a function invocation. For that reason, we also refer to internal requests
as application calls.

@@

All Argennon prefix trees have an equal branching factor \(\beta\). Therefore, we can represent an Argennon
prefix tree as a sequence of fractional numbers\footnote{It's possible to have \(a_i=0\). For
exmaple \(A^{(4)}=(0.2000)_{10}\) is correct.} in base \(\beta\):
\[
    (A^{(1)},A^{(2)},A^{(3)},\dots)\ ,
\]
where \(A^{(i)}=(0.a_{1}a_{2}\dots a_{i})_\beta\), and we have \(A^{(i)} \leq A^{(i+1)}\). A typical choice for
\(\beta\) could be \(2^8\).

@@

The AscEE uses \emph{authenticated message passing} for authorizing operations. In this method, every execution
session has a set of authenticated messages, and applications explicitly pass those messages for authorizing operations.
Authentication of messages can be efficiently done in parallel, using any type of cryptographic signature scheme in
the preprocessing phase of request execution. For example, we could use BLS aggregate signature to authenticate all
messages of a block in bulk.

Moreover, applications can use built-in functions of AscEE to generate authenticated messages in run-time. This enables
an application to authorize operations even for applications that are not directly called.

In addition to authenticated messages the AscEE provides a set of
cryptographic functions for validating signatures and cryptographic entities. By using these functionalities and
passing cryptographic signatures as parameters to methods, a programmer, having users' public keys, can implement
the required logic for authorizing operations.

Authorization by explicit messages and signatures eliminates the need for approval methods or call back patterns in
Argennon.

@@

Completing an execution session requires computational resources. The amount of resources used by an execution session
should be monitored, otherwise a malicious user would be able to easily spam and exhaust resources of the execution
environment. Resource usage can be measured per session or per application call.

The AscEE has two type of execution sessions: \emph{optimistic} and \emph{monitored}. Resource usage of an optimistic
session is always measured per session and default pre-defined resource caps are used. On the other hand in a monitored
session, some resources are measured per application call and some
caps are determined dynamically by the external HTTP request (i.e.\ transaction). An optimistic session must always
complete normally, a monitored session is allowed to either complete normally or abruptly.

Obviously resource monitoring is easier and more efficient for optimistic sessions. However, if
an optimistic session violates its resource caps, determining the point of failure requires precise resource
measurement. Note that for implementing optional decoupling principle, we need to determine the exact application
which has failed in a call chain. For example, assume that an optimistic session containing an application call
violates a 2 milliseconds execution time cap. If the call happens exactly after 2 milliseconds
in the caller application, a small fluctuation in the execution time measurement can
change the point of failure between the called and the main application. This can introduce nondeterministic
behaviour which could make block validation impossible. That's why monitored sessions exist. For a monitored session
this is not an issue because resource measurements and caps are defined per application call. This way, we can make
sure that resource caps are not too small and can be measured easily.

Different computational resources are measured and monitored during an AscEE session:
\begin{itemize}
    \item \textbf{execution time}:
    is the amount of cpu time that is required for executing a session or an application call. The execution time is
    measured in \emph{AscEE clocks}. One AscEE clock is defined as 1/1000 of the amount
    of \textbf{cpu time} needed for executing a predefined standard application which is used for benchmarking a host's
    performance.

    Optimistic sessions have a predefined \texttt{maxClocks} value which is determined by the Argennon protocol. This
    value defines a bound on the \textbf{total} cpu time of the session and no per application measurement is done.

    Monitored sessions perform per application call cpu-time measurement, and every application call during a monitored
    session has a separate \texttt{maxClocks} value. This value determines the maximum amount of time that the cpu
    can be used for executing that particular application call. It should be noted that the cpu timer of the
    application call is paused when the application makes a call to another application, and is resumed when the
    control returns. An application call
    needs at least 100 clocks and if the value of its \texttt{maxClock} is lower than this value the call will be
    considered a failed call.

    Each application call has some amount of \texttt{externalClocks}. When an application makes a request to another
    application it has to \emph{forward} a portion of its external clocks to the called application. This amount
    will determine the value of \texttt{maxClocks} for the called application.

    The amount of external clocks of an application
    call is defined to be \(2/3\) of its \texttt{maxClocks}. As a result, the total number of clocks of a monitored
    session is always less than \(3 \times \texttt{maxClocks}\) of the root application call. The value of
    \texttt{maxClocks} for
    the root application call is determined by the external request (transaction).

    \item \textbf{local memory}:
    any memory usage of an application that is not part of a heap chunk and is not part of another resource
    will be considered as local memory usage. Local memory is not persistent and when an application finishes serving a
    request and returns the HTTP response (i.e.\ the application call completes) its local memory is deleted.
    This resource is measured in bytes.

    Optimistic sessions measure local memory usage per session and enforce a protocol-defined cap on the total amount
    of local memory a session can use. Monitored sessions measure local memory usage per
    application call and enforce a protocol-defined cap per application call. An application call which tries to use
    more local memory than the cap, will fail.
    \item \textbf{heap access list}:
    every session can only access heap locations that are declared in its access list. In addition,
    resizing heap chunks can only be done in the range of the pre-declared lower bound and upper bound.
    \item \textbf{applications list}:
    a session may only make requests to applications that are declared in its application list.
    \item \textbf{call depth}:
    during a session the number of nested application calls can not be more than a threshold. This threshold is
    determined by the Argennon protocol.
    \item \textbf{active differed calls}:
    \item \textbf{virtual signatures}:
    \item \textbf{number of entrance locks}:
\end{itemize}

>>>>

An Argennon Standard Representation (ASR) is a programming language for describing argennon applications, optimized
for the architecture and requirements of the Argennon platform.
Argennon supports two standard representations: one is a high level text based language which needs
costly compilation before being executed on a hardware machine. The other is a low level binary representation which
usually can be executed, with minimal pre-processing, by a JIT compiler or an emulator. The high level
language is intended for preserving the high level information of applications logic to facilitate
platform specific compiler optimization at host nodes. The low level language, on the oder hand, is designed for
efficient direct execution of applications that are not frequently used.

The state of an Argennon application is stored in byte addressable finite arrays of memory called
\emph{heap chunks}. An application may have several heap chunks with different sizes, and can remove or
resize its heap chunks or allocate new chunks. Every chunk belongs to exactly one application and can only be modified
by its owner. In addition to heap chunks, every application has an amount of non-persistent local memory for
storing temporary data.

@@

The Argennon Smart Contract Execution Environment can be seen as a machine for executing Argennon applications to
fulfill HTTP requests, produce their HTTP responses and update related heap chunks. The execution of requests can be
considered sequential\footnote{Actually requests are executed in parallel but by performing data dependency analysis the
result is guaranteed to be identical
with sequential execution of requests.} and each request has a separate \emph{execution session}. An execution session
is a separate session of executing smart contract's code in order to fulfill an \emph{external} HTTP request.
External requests are requests that are not made by other Argennon applications.

@@

During an execution session an application can make \emph{internal} HTTP requests to other applications. Those
requests will not start a new execution session and will be executed within the current session. In AscEE making an
HTTP request to an application is similar to a function invocation, and for that reason, we also refer to them as
application calls.

@@

All Argennon prefix trees have an equal branching factor \(\beta\). Therefore, we can represent an Argennon
prefix tree as a sequence of fractional numbers\footnote{It's possible to have \(a_i=0\).
For example, \(A^{(4)}=(0.2000)_{10}\) is correct.} in base \(\beta\):
\[
    (A^{(1)},A^{(2)},A^{(3)},\dots)\ ,
\]
where \(A^{(i)}=(0.a_{1}a_{2}\dots a_{i})_\beta\), and we have \(A^{(i)} \leq A^{(i+1)}\).\footnote{A typical choice for
    \(\beta\) is \(2^8\).}

@@

The AscEE uses \emph{Authenticated Message Passing} for authorizing operations. In this method, every execution
session has a set of authenticated messages, and applications \textbf{explicitly} pass those messages for authorizing
operations.
Authentication of messages can be efficiently done in parallel, using any type of cryptographic signature scheme in
the preprocessing phase of request execution. For example, we could use BLS aggregate signature to authenticate all
messages of a block in bulk.

Moreover, applications use built-in functions of AscEE to generate authenticated messages in run-time. This enables
an application to authorize operations for another application even if that application is not directly called by
itself.

In addition to authenticated messages, the AscEE provides a set of
cryptographic functions for validating signatures and cryptographic entities. By using these functions and
passing cryptographic signatures as parameters to methods, a programmer, having users' public keys, can implement
the required logic for authorizing operations.

Authorization by explicit authenticated messages and signatures eliminates the need for approval mechanisms or call back
patterns in Argennon.

@@

Completing an execution session requires computational resources. The amount of resources used by an execution session
should be monitored and managed, otherwise a malicious user would be able to easily spam and exhaust resources of the
execution environment.

Resource usage can be measured per session or per application call. Obviously per session measurement is
easier and more efficient. However, when we measure resources per session if
a session violates its resource caps, determining the point of failure requires precise and error-free resource
measurement. Note that for implementing optional decoupling principle, we need to determine the exact application call
which has failed in a call chain. For example, assume that a session containing an application call
violates a 2 milliseconds execution time cap. If the call happens exactly after 2 milliseconds
in the caller's code, a small fluctuation in the execution time measurement can
change the point of failure between the called and the main application. This can introduce nondeterministic
behaviour which could make block validation impossible. That's why the AscEE sometimes uses per application call
resource management. If we define the resource caps per application call and perform our measurements for each
application call separately, by ensuring that caps are larger than the measurement error, errors in the
measurements would not change the point of failure in the call chain anymore.

The AscEE has two type of execution sessions: \emph{optimistic} and \emph{monitored}. Resource usage of an optimistic
session is always measured per session and default pre-defined resource caps are used. On the other hand, for a
monitored session, resources that can not be measured precisely are measured per application call and their caps are
determined per application call by the external HTTP request (i.e.\ transaction). An optimistic session must
always complete normally, a monitored session is allowed to either complete normally or abruptly.

Different computational resources are measured and monitored during an AscEE session:
\begin{itemize}
    \item \textbf{execution time}:
    is the amount of cpu time that is required for executing a session or an application call. The execution time is
    measured in \emph{AscEE clocks}. One AscEE clock is defined as 1/1000 of the amount
    of \textbf{cpu time} needed for executing a predefined standard application which is used for benchmarking a host's
    performance. In other words, by definition the AscEE imaginary standard machine completes the standard benchmark
    in 1000 clocks.

    Optimistic sessions have a predefined \texttt{maxClocks} value which is determined by the Argennon protocol. This
    value defines a bound on the \textbf{total} AscEE clocks of the session, and no per application measurement is done.

    Monitored sessions perform per application call cpu-time measurement, and every application call during a monitored
    session has a separate \texttt{maxClocks} value. This value determines the maximum amount of time that the cpu
    can be used for executing that particular application call. It should be noted that the cpu timer of the
    application call is paused when the application makes a call to another application, and is resumed when the
    control returns. An application call
    needs at least 10 clocks and if the value of its \texttt{maxClock} is lower than this value the call will be
    considered a failed call.

    Each application call has some amount of \texttt{externalClocks}. When an application makes a request to another
    application it has to \emph{forward} a portion of its \textbf{external} clocks to the called application. This
    amount will determine the value of \texttt{maxClocks} for the called application.

    The amount of external clocks of an application
    call is defined to be \(2/3\) of its \texttt{maxClocks}. As a result, the total number of clocks of a monitored
    session is always less than \(3 \times \texttt{maxClocks}\) of the root application call. The value of
    \texttt{maxClocks} for
    the root application call is determined by the external request (i.e.\ transaction).

    \item \textbf{local memory}:
    any memory usage of an application that is not part of a heap chunk and is not part of another measured resource
    will be considered as local memory usage. Local memory is not persistent and when an application finishes serving a
    request and returns the HTTP response (i.e.\ the application call completes) its local memory is deleted.
    This resource is measured in bytes.

    Optimistic sessions measure local memory usage per session and enforce a protocol-defined cap on the total amount
    of local memory a session can use. Monitored sessions measure local memory usage per
    application call and enforce a protocol-defined cap for each application call separately. An application call
    which tries to use more local memory than the cap, fails.
    \item \textbf{heap access list}:
    every session can only access heap locations that are declared in its access list. In addition,
    resizing heap chunks can only be done in the range of the pre-declared lower bound and upper bound.
    \item \textbf{app access list}:
    a session may only make requests to applications that are declared in its application access list.
    \item \textbf{call depth}:
    during a session the number of nested application calls can not be more than a threshold. This threshold is
    determined by the Argennon protocol. It should be noted that a differed call is considered like a normal call and
    increases the call depth by one level.
    \item \textbf{differed calls}:
    every application call can have a limited number of differed calls which is determined by the protocol.
    To simplify the implementation, this limit is defined per application call. Since in Argennon the call depth is
    limited, a per-application call limit will also define an implicit limit for the total number of active differed
    calls.
    \item \textbf{virtual signatures}:
    \item \textbf{number of entrance locks}:
\end{itemize}
